---
title: "R Notebook"
output: html_notebook
---

# Big Data Analysis
## In this project, we try to explore the relationship between British-American Tobacco's stock prices and google trends.
## We will be using 3 keywords to explore the relationship. i.e. 
###1. Quit Smoking
###2. Effects of smoking
###3. Electronic cigarette
```{r}
# Importing libraries
library(ggplot2)
library(tidyr)
library(dplyr)
library(scales)
library(reshape2)
library(lubridate)
library(gridExtra)
library(readxl)
library(GGally)
library(cowplot)
library(plotly)
library(corrplot)
library(forecast)
library(grid)
library(caret)
```

```{r}
# Reading Stock Price data
stock_data <- tbl_df(read.csv("BTC.csv"))

```

```{r}
# Extracting year and month
stock_data <-
  mutate(stock_data, year_month = format(as.Date(stock_data$date, format =
                                                "%m/%d/%Y"), "%Y-%m"))

```

```{r}
# Calculating average stock price monthly

stock_data_by_month <- stock_data %>%
  group_by(year_month) %>%
  summarise(
  stock_price = mean(close)
  ) %>%
  arrange(year_month)

```


```{r}
# Reading google trends data
key1 <- tbl_df(read.csv("quitsmoking.csv"))
key2 <- tbl_df(read.csv("EofSmoking.csv"))
key3 <- tbl_df(read.csv("ECig.csv"))

colnames(key1)[1]<-"year_month"
colnames(key1)[2]<-"quitsmoking"
  
colnames(key2)[1]<-"year_month"
colnames(key2)[2]<-"effectsofsmoking"

colnames(key3)[1]<-"year_month"
colnames(key3)[2]<-"electroniccigarette"

```


```{r}
# Merging all data
allData <-merge(key1,stock_data_by_month,id="year_month")
allData<-merge(key2,allData,id="year_month")
allData<-merge(key3,allData,id="year_month")

```



```{r}
allData <-
  mutate(allData, date = paste0(allData$year_month,"-01"))
allData$date<-as.Date(allData$date)
```


# Summary of all Data
```{r}
# Checking out the summary of data
summary(allData)
```

```{r}
# Helper function to plot multiple graphs
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

# Data distribution
## We use density plots to observe the data distribution
```{r}
p1<-ggplot(data=allData, aes(allData$stock_price )) + 
  geom_histogram(aes(y=..density..),col="red", 
                 fill="green", 
                 alpha = .2)+geom_density(col=2)+xlab("Stock price (£)")+ggtitle("A")


p2<-ggplot(data=allData, aes(allData$quitsmoking )) + 
  geom_histogram(aes(y=..density..),col="red", 
                 fill="green", 
                 alpha = .2)+geom_density(col=2) + xlab("Quit smoking (GSV)")+ggtitle("B")

p3<-ggplot(data=allData, aes(allData$effectsofsmoking)) + 
  geom_histogram(aes(y=..density..),col="red", 
                 fill="green", 
                 alpha = .2)+geom_density(col=2) +xlab("Effects of smoking (GSV)")+ggtitle("C")

p4<-ggplot(data=allData, aes(allData$electroniccigarette )) + 
  geom_histogram(aes(y=..density..),col="red", 
                 fill="green", 
                 alpha = .2)+geom_density(col=2) + xlab("Electronic ciggarette (GSV)")+ggtitle("D")
multiplot(p1, p2, p3, p4, cols=2)

```



# Plotting monthly values of Stock price against google trends

```{r,fig.width = 16, fig.height = 5}
p <- ggplot(allData, aes(x = date))
  p <- p + geom_line(aes(y = stock_price, colour = "Stock Price",group="stock_price"))
  
  # adding the relative humidity data, transformed to match roughly the range of the temperature
  p <- p + geom_line(aes(y = quitsmoking, colour = "Quit smoking",group="quitsmoking"))
  
  # now adding the secondary axis, following the example in the help file ?scale_y_continuous
  # and, very important, reverting the above transformation
  p <- p + scale_y_continuous(sec.axis = sec_axis(~.*1, name = "Google Search"))
  
  # modifying colours and theme options
  p <- p + scale_colour_manual(values = c("blue", "red"))
  p <- p + labs(y = "Stock Price (£)",
                x = "Months",
                colour = " ")
  p1 <- p + theme(legend.position = c(0.8, 0.9))+ggtitle("A. Stock price vs Quit Smoking")+theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+theme(axis.text = element_text(size=12),
          axis.title=element_text(size=16))

```



```{r,fig.width = 16, fig.height = 5}
p <- ggplot(allData, aes(x = date))
  p <- p + geom_line(aes(y = stock_price, colour = "Stock Price",group="stock_price"))
  
  # adding the relative humidity data, transformed to match roughly the range of the temperature
  p <- p + geom_line(aes(y = effectsofsmoking, colour = "Effects of smoking",group="effectsofsmoking"))
  
  # now adding the secondary axis, following the example in the help file ?scale_y_continuous
  # and, very important, reverting the above transformation
  p <- p + scale_y_continuous(sec.axis = sec_axis(~.*1, name = "Google Search"))
  
  # modifying colours and theme options
  p <- p + scale_colour_manual(values = c("blue", "red"))
  p <- p + labs(y = "Stock Price (£)",
                x = "Months",
                colour = " ")
  p2 <- p + theme(legend.position = c(0.8, 0.9))+ggtitle("B. Stock price vs Effects of Smoking")+theme_minimal() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())+theme(axis.text = element_text(size=12),
          axis.title=element_text(size=16))

```


```{r,fig.width = 16, fig.height = 5}

p <- ggplot(allData, aes(x = date))
  p <- p + geom_line(aes(y = stock_price, colour = "Stock Price",group="stock_price"))
  
  # adding the relative humidity data, transformed to match roughly the range of the temperature
  p <- p + geom_line(aes(y = electroniccigarette, colour = "Electronic cigarette",group="electroniccigarette"))
  
  # now adding the secondary axis, following the example in the help file ?scale_y_continuous
  # and, very important, reverting the above transformation
  p <- p + scale_y_continuous(sec.axis = sec_axis(~.*1, name = "Google Search"))
  
  # modifying colours and theme options
  p <- p + scale_colour_manual(values = c("blue", "red"))
  p <- p + labs(y = "Stock Price (£)",
                x = "date",
                colour = " ")
  p3 <- p + theme(legend.position = c(0.8, 0.9))+ggtitle("C. Stock price vs Electronic cigarette")+theme_minimal() +theme(axis.text.x = element_text(angle = 90))+scale_x_date(date_breaks = "8 month")+theme(axis.text = element_text(size=10),
          axis.title=element_text(size=16))
  


```

```{r,fig.width = 12, fig.height = 8}
grid.arrange(p1,p2,p3)
```


```{r}
# Standardizing the data

allData$quitsmoking_std<-as.numeric(scale(allData$quitsmoking ))
allData$effectsofsmoking_std<-as.numeric(scale(allData$effectsofsmoking ))
allData$electroniccigarette_std<-as.numeric(scale(allData$electroniccigarette ))
allData$stock_price_std<-as.numeric(scale(allData$stock_price))

```


# Normality test for stock prices
```{r}
shapiro.test(allData$stock_price_std)
```


# Observing co-relation
## Since data-distributions do not seem to be normal, we will use kendall's co-relation test
## 1. Effects of smoking

###Normality test
```{r}
shapiro.test(allData$effectsofsmoking_std )
```

```{r}
cor.test(allData$effectsofsmoking ,allData$stock_price_std,method="kendall")

```

```{r}
linearModel<-lm(effectsofsmoking_std~stock_price_std,data=allData)
summary(linearModel)
```


```{r}
ggplot(allData, aes(x=effectsofsmoking_std , y=stock_price_std)) + geom_point()+geom_smooth(method=lm)+xlab("Effects of smoking(normalised)")+ylab("Stock price(normalised)")+ggtitle("Scatter plot of Google trend vs Stock price")

```


## 2. Quit smoking
###Normality test
```{r}
shapiro.test(allData$quitsmoking_std)
```

```{r}
cor.test(allData$quitsmoking_std ,allData$stock_price_std,method="kendall")

```

```{r}
linearModel<-lm(quitsmoking_std~stock_price_std,data=allData)
summary(linearModel)
```


```{r}
ggplot(allData, aes(x=quitsmoking_std , y=stock_price_std)) + geom_point()+geom_smooth(method=lm)+xlab("Quit smoking(normalised)")+ylab("Stock price(normalised)")+ggtitle("Scatter plot of Google trend vs Stock price")

```





## 3. Electronic-Cigarrette
###Normality test
```{r}
shapiro.test(allData$electroniccigarette_std)
```

```{r}
cor.test(allData$electroniccigarette_std,allData$stock_price_std,method="kendall")
```




```{r}
linearModel<-lm(electroniccigarette_std~stock_price_std,data=allData)
summary(linearModel)
```



```{r}
ggplot(allData, aes(x= electroniccigarette_std, y=stock_price_std)) + geom_point()+geom_smooth(method=lm)+xlab("Electronic Cigarette(normalised)")+ylab("Stock price(normalised)")+ggtitle("Scatter plot of Google trend vs Stock price")
```


```{r,fig.height=5}
library(corrplot)
cor_matrix<-cor(allData[,c("stock_price","quitsmoking","electroniccigarette","effectsofsmoking")])
corrplot(cor_matrix,method="number")

```





# Dividing data into training and test set
## We will use 90% of the data as train, and 10% as test
```{r}
ntest <- 9
len<-nrow(allData)
train<-allData[1:(len-ntest),]
test<-allData[(len-ntest+1):len,]

```


```{r,fig.width = 10, fig.height = 5}
# Making TS objects
stock_price_train<- ts(train$stock_price_std, frequency=12)
ecigg_train<-ts(train$electroniccigarette_std ,frequency = 12)
effect_train<-ts(train$effectsofsmoking_std,frequency = 12)
quit_train <- ts(train$quitsmoking_std,frequency = 12)


stock_price_test<- ts(test$stock_price_std, frequency=12)
ecigg_test<-ts(test$electroniccigarette_std ,frequency = 12)
effect_test<-ts(test$effectsofsmoking_std,frequency = 12)
quit_test <- ts(test$quitsmoking_std,frequency = 12)


```



# Using Auto-Arima on training set (without regressors)
```{r}
model<-auto.arima(stock_price_train)
pred<-forecast(model,h=9)
plot(forecast(model,h=9))

```

```{r}
# predicted values of ARIMA
auto_arima_pred<-as.numeric(pred$mean)
without_reg<-as.numeric(pred$mean)
```


# Evaluating on test set
```{r}

postResample(auto_arima_pred,test$stock_price_std)
```




# Using Auto-Arima on training set (with regressors)
#1. Using E-cig as a regressor
```{r}
model<-auto.arima(stock_price_train,xreg=ecigg_train)
pred<-forecast(model,h=9,xreg=ecigg_test)
plot(pred)

```




```{r}
# predicted values of ARIMA
auto_arima_pred<-as.numeric(pred$mean)

```


# Evaluating on test set
```{r}

postResample(auto_arima_pred,test$stock_price_std)
```




# Using Auto-Arima on training set (with regressors)
#2. Using Quit Smoking as a regressor
```{r}
model<-auto.arima(stock_price_train,xreg=quit_train)
pred<-forecast(model,h=9,xreg=quit_test)
plot(pred)

```





```{r}
# predicted values of ARIMA
auto_arima_pred<-as.numeric(pred$mean)
```


# Evaluating on test set
```{r}

postResample(auto_arima_pred,test$stock_price_std)
```





# Using Auto-Arima on training set (with regressors)
#3. Using effects of smoking as a regressor
```{r}
model<-auto.arima(stock_price_train,xreg=effect_train)
pred<-forecast(model,h=9,xreg=effect_test)
plot(pred)

```




```{r}
# predicted values of ARIMA
auto_arima_pred<-as.numeric(pred$mean)
```


# Evaluating on test set
```{r}

postResample(auto_arima_pred,test$stock_price_std)
```





# Using Auto-Arima on training set (with regressors)
#4. Using all keywords as a regressor
```{r}
xreg=cbind(ecigg_train,quit_train,effect_train)
model<-auto.arima(stock_price_train,xreg=xreg)
xreg=cbind(ecigg_test,quit_test,effect_test)
pred<-forecast(model,h=9,xreg=xreg)
plot(pred)

```



```{r}
# predicted values of ARIMA
auto_arima_pred<-as.numeric(pred$mean)
with_reg<-as.numeric(pred$mean)
```


# Evaluating on test set
```{r}

postResample(auto_arima_pred,test$stock_price_std)
```



# Now Testing with lagged values
```{r}

# Dividing data into training and test set
## We will use 90% of the data as train, and 10% as test
ntest <- 9
len<-nrow(allData)
train<-allData[1:(len-ntest),]
test<-allData[(len-ntest+1):len,]

```


```{r,fig.width = 10, fig.height = 5}
# Making TS objects
stock_price_train<- ts(allData$stock_price_std, frequency=12)
ecigg_train<-ts(allData$electroniccigarette_std ,frequency = 12)
effect_train<-ts(allData$effectsofsmoking_std,frequency = 12)
quit_train <- ts(allData$quitsmoking_std,frequency = 12)


stock_price_train_lagged<-stock_price_train[2:87]
ecigg_train_lagged<-ecigg_train[1:86]
quit_train_lagged<-quit_train[1:86]
effect_train_lagged<-effect_train[1:86]

stock_price_test_lagged<-stock_price_train[88:96]
ecigg_test_lagged<-ecigg_train[87:95]
quit_test_lagged<-quit_train[87:95]
effect_test_lagged<-effect_train[87:95]
```




# Using Auto-Arima on training set (without regressors)
```{r}
model<-auto.arima(stock_price_train_lagged)
pred<-forecast(model,h=9)
plot(forecast(model,h=9))

```

```{r}
# predicted values of ARIMA
auto_arima_pred<-as.numeric(pred$mean)
without_reg_lag<-as.numeric(pred$mean)

```


# Evaluating on test set
```{r}

postResample(auto_arima_pred,test$stock_price_std)
```





# Using Auto-Arima on training set (with regressors)
#1. Using E-cig as a regressor


```{r}
model<-auto.arima(stock_price_train_lagged,xreg=ecigg_train_lagged)
pred<-forecast(model,h=9,xreg =ecigg_test_lagged )
plot(pred)

```

```{r}
# predicted values of ARIMA
auto_arima_pred<-as.numeric(pred$mean)
```


# Evaluating on test set
```{r}

postResample(auto_arima_pred,test$stock_price_std)
```



#2. Using quit-smoking as a regressor


```{r}
model<-auto.arima(stock_price_train_lagged,xreg=quit_train_lagged)
pred<-forecast(model,h=9,xreg =quit_test_lagged )
plot(pred)

```

```{r}
# predicted values of ARIMA
auto_arima_pred<-as.numeric(pred$mean)
```


# Evaluating on test set
```{r}

postResample(auto_arima_pred,test$stock_price_std)
```





#3. Using effect of smoking as a regressor


```{r}
model<-auto.arima(stock_price_train_lagged,xreg=effect_train_lagged)
pred<-forecast(model,h=9,xreg =effect_test_lagged )
plot(pred)

```

```{r}
# predicted values of ARIMA
auto_arima_pred<-as.numeric(pred$mean)
```


# Evaluating on test set
```{r}

postResample(auto_arima_pred,test$stock_price_std)
```








#4. Using all keywords as a regressor


```{r}
xreg=cbind(effect_train_lagged,quit_train_lagged,ecigg_train_lagged)
model<-auto.arima(stock_price_train_lagged,xreg=xreg)
xreg=cbind(effect_test_lagged,quit_test_lagged,ecigg_test_lagged)
pred<-forecast(model,h=9,xreg =xreg )
plot(pred)

```

```{r}
# predicted values of ARIMA
auto_arima_pred<-as.numeric(pred$mean)
with_reg_lag<-as.numeric(pred$mean)

```


# Evaluating on test set
```{r}

postResample(auto_arima_pred,test$stock_price_std)
```



```{r}
# Creating datafame with predicted stock prices of last 9 months with diff models
temp<-data.frame(date=allData[88:96,"year_month"],original_stock_price=allData[88:96,"stock_price_std"],predicted_with_regressor=with_reg,predicted_without_regressor=without_reg,predicted_with_regressor_lag=with_reg_lag,predicted_without_regressor_lag=without_reg_lag)

```


# Comparing Original and Predicted stock prices 
```{r,fig.width=10}
p <- ggplot(temp, aes(x = date))
  # adding the relative humidity data, transformed to match roughly the range of the temperature
  p <- p + geom_line(aes(y = original_stock_price, colour = "Original Stock price",group="original_stock_price"))
    p <- p + geom_line(aes(y =predicted_with_regressor, colour = "Prediced with regressor",group="predicted_with_regressor"))
  p <- p + geom_line(aes(y = predicted_without_regressor, colour = "Prediced without regressor",group="predicted_without_regressor"))
  p <- p + geom_line(aes(y = predicted_with_regressor_lag, colour = "Prediced with regressor and lagged values",group="predicted_with_regressor_lag"))
  p <- p + geom_line(aes(y = predicted_without_regressor_lag, colour = "Predicted without regressor and lagged values",group="predicted_without_regressor_lag"))

  # now adding the secondary axis, following the example in the help file ?scale_y_continuous
  # and, very important, reverting the above transformation
  
  # modifying colours and theme options
  p <- p + scale_colour_manual(values = c("red", "blue","green","purple","orange"))
  p <- p + labs(y = "Stock Price",
                x = "Months",
                colour = " ")
  p <- p + theme(legend.position = c(0.8, 0.9))+ggtitle("Original vs Predicted stock price")+theme_minimal()+ theme(axis.text.x = element_text(angle = 90))+theme(axis.text = element_text(size=12),
          axis.title=element_text(size=16))
p 
```



